---
title: "Positive-control primer efficiency (4 housekeeping/pluripotency primers, Maria gDNA, iTaq SYBR)"
author: "Wonhyung Choi"
date: "2026-07-16"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: kable
---
```{r libs}
library(ggplot2)
library(pals)
```

# Config

Four candidate positive-control primers (Hsp90aa, gapdh, klf4, nanog) titrated over
1 / 10 / 100 ng gDNA (n = 2). Maria_GAPDH run alongside as the established reference PC.

```{r config}
BASE_DIR <- path.expand("~/Desktop/Domcke Lab/Master Thesis/Results/cell_type_specific_primer_test")
RESULTS_DIR <- file.path(BASE_DIR, "20260716_PC_efficiency")
CSV_PATH <- file.path(RESULTS_DIR, "admin_2026-07-16 14-31-44_782BR21735.csv")

pal <- trubetskoy(5)
names(pal) <- NULL
pal <- c(
  "Hsp90aa" = pal[1],
  "gapdh" = pal[2],
  "klf4" = pal[3],
  "nanog" = pal[4],
  "Maria_GAPDH" = pal[5]
)

dir.create(RESULTS_DIR, recursive = TRUE, showWarnings = FALSE)
```

# 1. Read data

The metadata header is 19 lines; the well table starts on line 20.

```{r read}
dat <- read.csv(CSV_PATH, skip = 19, header = TRUE, stringsAsFactors = FALSE)
dat <- dat[, c("Well", "Target", "Content", "Sample", "Cq", "Melt.Temperature")]

dat$Cq <- as.numeric(dat$Cq)                 # "NaN" strings -> NA
dat$Melt.Temperature <- as.numeric(dat$Melt.Temperature)  # "None" strings -> NA
dat$ng <- as.numeric(sub("ng", "", dat$Sample))
dat$log_ng <- log10(dat$ng)
dat
```

# 2. Linear fit and efficiency per target

Linear fit `Cq ~ log10(ng)`; efficiency (%) = `(10^(-1/slope) - 1) * 100`.

```{r fit}
hsp90aa <- dat[dat$Target == "Hsp90aa", ]
gapdh   <- dat[dat$Target == "gapdh", ]
klf4    <- dat[dat$Target == "klf4", ]
nanog   <- dat[dat$Target == "nanog", ]
mgapdh  <- dat[dat$Target == "Maria_GAPDH", ]

fit_hsp90aa <- lm(Cq ~ log_ng, data = hsp90aa)
fit_gapdh   <- lm(Cq ~ log_ng, data = gapdh)
fit_klf4    <- lm(Cq ~ log_ng, data = klf4)
fit_nanog   <- lm(Cq ~ log_ng, data = nanog)
fit_mgapdh  <- lm(Cq ~ log_ng, data = mgapdh)

eff_hsp90aa <- (10^(-1 / coef(fit_hsp90aa)[2]) - 1) * 100
eff_gapdh   <- (10^(-1 / coef(fit_gapdh)[2])   - 1) * 100
eff_klf4    <- (10^(-1 / coef(fit_klf4)[2])    - 1) * 100
eff_nanog   <- (10^(-1 / coef(fit_nanog)[2])   - 1) * 100
eff_mgapdh  <- (10^(-1 / coef(fit_mgapdh)[2])  - 1) * 100

eff_tbl <- data.frame(
  Target = c("Hsp90aa", "gapdh", "klf4", "nanog", "Maria_GAPDH"),
  slope = c(coef(fit_hsp90aa)[2], coef(fit_gapdh)[2], coef(fit_klf4)[2],
            coef(fit_nanog)[2], coef(fit_mgapdh)[2]),
  R2 = c(summary(fit_hsp90aa)$r.squared, summary(fit_gapdh)$r.squared,
         summary(fit_klf4)$r.squared, summary(fit_nanog)$r.squared,
         summary(fit_mgapdh)$r.squared),
  Efficiency = c(eff_hsp90aa, eff_gapdh, eff_klf4, eff_nanog, eff_mgapdh),
  MeltTm = c(mean(hsp90aa$Melt.Temperature, na.rm = TRUE),
             mean(gapdh$Melt.Temperature, na.rm = TRUE),
             mean(klf4$Melt.Temperature, na.rm = TRUE),
             mean(nanog$Melt.Temperature, na.rm = TRUE),
             mean(mgapdh$Melt.Temperature, na.rm = TRUE))
)
rownames(eff_tbl) <- NULL
eff_tbl
```

# 3. Standard curves

Colour = primer. Efficiency close to 100 % and a clean single melt peak indicate a usable
positive control.

```{r plot, fig.cap="Cq vs log10 gDNA input for the four candidate positive-control primers, with Maria_GAPDH as reference."}
plot_df <- dat[!is.na(dat$Cq), ]
plot_df$Target <- factor(plot_df$Target,
  levels = c("Hsp90aa", "gapdh", "klf4", "nanog", "Maria_GAPDH"))

p_std <- ggplot(plot_df, aes(x = log_ng, y = Cq, colour = Target)) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.6) +
  geom_point(size = 2, alpha = 0.9) +
  scale_colour_manual(values = pal, name = "Primer") +
  labs(x = "log10(gDNA input, ng)", y = "Cq",
       title = "Positive-control primer standard curves (Maria gDNA, iTaq SYBR)") +
  theme_bw(base_size = 10)
p_std

ggsave(file.path(RESULTS_DIR, "standard_curves_20260716.png"), p_std, width = 7.6, height = 5.4, dpi = 150)
ggsave(file.path(RESULTS_DIR, "standard_curves_20260716.pdf"), p_std, width = 7.6, height = 5.4)
```
